The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-01-27 src/ms.tex
Add first version of CTA section

Cherenkov Telescope Array
\secauthor{Axel Donath, Max Noethe}

@@

{\it Learn Astropy} is an umbrella term that acknowledges the broad educational efforts made by the Astropy Project, which are led by the Learn Team.
The efforts focus on developing online content and workshops covering astro-specific coding tasks in Python.
As described in \textsl{[Astropy-ii paper]}, There are four different types of Learn Astropy content: \textit{tutorials}, consisting of Jupyter Notebook lessons that are published in HTML format online; \textit{guides}, which are a series of lessons providing a foundational resource for performing a certain type of astronomical analyses; \textit{examples}, which are snippets of code that showcase a short task that can be performed with Astropy or an affiliated package; and \textit{documentation}, which is contained within the code base.
This categorization drives content development, infrastructure choices, and the appearance of the Learn Astropy website.
The Learn Team meets weekly to work on creating, expanding, improving these educational resources.

@@

\item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
be on a paid plan; additionally NumFOCUS has negotiated a special rate for
open-source projets. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easily-to-navigate interface for users to ask for help Accordingly, we have commissioned
a Discourse site which is more open than the Facebook group and more user-oriented
than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
developers are able to invoice as independent contractors the time they
spend helping users on these forums.

\secunfilled

>>>>

Cherenkov Telescope Array
\secauthor{Axel Donath, Maximilian Nöthe}

The Cherenkov Telescope (CTA) will be the next generation very-high-energy
gamma-ray observatory.
It will improve over the current generation of imaging atmospheric Cherenkov telescopes (IACTs)
by a factor of five to ten in sensitivity and it will be able to observe the whole sky from a combination of two sites:
a northern site in La Palma, Spain, and a southern one in Paranal, Chile.
CTA will be able to observe gamma rays in the energy range between $20\,\mathrm{GeV}$ and $300\,\mathrm{GeV}$
using three different types of telescopes, in total over 100 telescopes are planned at the two sites.
CTA will also be the first open gamma-ray observatory.

The data analysis pipeline is developed as open source software and essentially split in two domains:
\begin{enumerate}
  \item In the low-level analysis, the properties of the recorded air shower events
    have to be estimated from the raw data.
    The raw data consists of very short (\textasciitilde $40\,\mathrm{ns}$) videos recorded with the fast and
    sensitive cameras of the telescopes.
    This includes the energy, particle type and direction of origin of the particle that induced the air shower
    and the time the shower was recorded.
  \item In the higher-level analysis, these reconstructed event lists are used together with the
    characterization of the measurement process and the low-level analysis to perform the actual scientific analysis.
    This includes energy spectra, light curves, sky maps and many more.
\end{enumerate}

A prototype for the low-level analysis is ctapipe~\cite{ctapipe},
a python package developed to perform all the necessary tasks to from the raw data
of Cherenkov telescopes to the reconstructed event lists.

The high-level analysis will be based on the Gammapy~\cite{gammapy} library, which
will be the core of the CTA science tools.

Both Gammapy and \texttt{ctapipe} are using astropy heavily, mainly for units, coordinate transformations,
tables and FITS IO.

As CTA will record gamma-ray events with a rate of up to $10\,000$ events per second,
it needs to perform a large number of coordinate transformations.
To enable this, CTA member M.~Nöthe contributed a major performance improvement
for large numbers of coordinates with different observation times,
based on earlier work by B.~Winkel.

Together with Gammapy maintainer A.~Donath and former maintainer C.~Deil,
a total of 107 merged pull requests were contributed to ctapipe.

@@

{\it Learn Astropy} is an umbrella term that acknowledges the broad educational
efforts made by the Astropy Project, which are led by the Learn Team.
The efforts focus on developing online content and workshops covering
astro-specific coding tasks in Python.
As described in \citet{astropy:2018}, There are four different types of
Learn Astropy content: \textit{tutorials}, consisting of Jupyter Notebook
lessons that are published in HTML format online; \textit{guides}, which are a
series of lessons providing a foundational resource for performing certain
type of astronomical analyses; \textit{examples}, which are snippets of code
that showcase a short task that can be performed with Astropy or an affiliated
package; and \textit{documentation}, which is contained within the code base.
This categorization drives content development, infrastructure choices, and the
appearance of the Learn Astropy website.
The Learn Team meets weekly to work on creating, expanding, improving these
educational resources.

@@

\begin{itemize}
\item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
be on a paid plan; additionally NumFOCUS has negotiated a special rate for
open-source projets. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easily-to-navigate interface for users to ask for help Accordingly, we have commissioned
a Discourse site which is more open than the Facebook group and more user-oriented
than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
developers are able to invoice as independent contractors the time they
spend helping users on these forums.
\end{itemize}
\secunfilled
